---
output:
  md_document:
    variant: gfm
---

```{r, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

# bettermc

[![R build status](https://github.com/gfkse/bettermc/workflows/R-CMD-check/badge.svg)](https://github.com/gfkse/bettermc/actions?workflow=R-CMD-check)
[![codecov](https://codecov.io/gh/gfkse/bettermc/branch/master/graph/badge.svg?token=FYYM156COF)](https://codecov.io/gh/gfkse/bettermc)

The `bettermc` package provides a wrapper around the `parallel::mclapply` function for better performance, error handling and UX. 

**Note:** As of now this package only works on Linux, not macOS. Windows will never be supported as `parallel::mclapply` itself does not support it.

## Features
Here is a short overview on its main features ...

### Traceback and Crashdumps
By default, crashdumps and full tracebacks are generated on errors in child processes:
``` r
g <- function(x) x + 1
f <- function(x) g(as.character(x))
bettermc::mclapply(1:2, f)
```

    ## Error in bettermc::mclapply(1:2, f) :
    ##   error(s) occured during mclapply; first original message:

    ## Error: non-numeric argument to binary operator

    ## Traceback:
    ## 3: g(as.character(x))
    ## 2: (function (x) 
    ##    g(as.character(x)))(1L)
    ## 1: bettermc::mclapply(1:2, f)

    ## Crash dump avilable. Use 'debugger(attr(*, "dump.frames"))' for debugging.

``` r
# in a non-interactive session a file "last.dump.rds" is created
last.dump <- readRDS("last.dump.rds")
debugger(attr(last.dump[[1L]], "dump.frames"))
```

    ## Message:  non-numeric argument to binary operatorAvailable environments had calls:
    ## 1: (function (x) 
    ## g(as.character(x)))(1)
    ## 2: #1: g(as.character(x))

    ## Enter an environment number, or 0 to exit  
    ## Selection: 0

### Output, Messages and Warnings
In contrast to `parallel::mclapply`, neither output nor messages nor warnings from the child processes are lost.
All of these can be forwarded to the parent process and are prefix with the index of the function invocation from which they originate:
```{r}
f <- function(i) {
  if (i == 1) message(letters[i])
  else if (i == 2) warning(letters[i])
  else print(letters[i])
  
  i
}
ret <- bettermc::mclapply(1:4, f)
```

### POSIX Shared Memory
Many types of objects can be returned from the child processes using POSIX shared memory.
This includes e.g. logical, numeric, complex and raw vectors and arrays as well as factors.
In doing so, the overhead of getting larger results back into the parent R process is reduced:
```{r}
X <- data.frame(
  x = runif(3e7),
  y = sample(c(TRUE, FALSE), 3e7, TRUE),
  z = 1:3e7
)
f <- function(i) X

microbenchmark::microbenchmark(
  bettermc1 = bettermc::mclapply(1:2, f, mc.share.copy = FALSE),
  bettermc2 = bettermc::mclapply(1:2, f),
  bettermc3 = bettermc::mclapply(1:2, f, mc.share.vectors = FALSE),
  bettermc4 = bettermc::mclapply(1:2, f, mc.share.vectors = FALSE, mc.shm.ipc = FALSE),
  parallel = parallel::mclapply(1:2, f),
  times = 10, setup = gc()
)
```

In examples `bettermc1` and `bettermc2`, the child processes place the columns of the return value `X` in shared memory. 
The object which needs to be serialized for transfer from child to parent processes hence becomes:
```{r}
X_shm <- bettermc:::vectors2shm(X, name_prefix = "/bettermc_README_")
str(X_shm)
```

Column `z` is an `ALTREP` and, because it can be serialized efficiently, is left alone by default. 
The parent process can recover the original object from `X_shm`:
```{r}
Y <- bettermc:::shm2vectors(X_shm)
stopifnot(identical(X, Y))
```

The internal functions `vectors2shm()` and `shm2vectors()` recursively walk the return value and apply the exported functions `copy2shm()` and `allocate_from_shm()`, respectively.

In `bettermc1`, the shared memory objects are used directly by the parent process. 
In `bettermc2`, which is the default, new vectors are allocated in the parent process and the data is merely copied from the shared memory objects, which are freed afterwards. See `?copy2shm` for more details on this topic and why the slower `mc.share.copy = TRUE` might be a sensible default.

In `bettermc3`, the original `X` is serialized and the resulting raw vector is placed in shared memory, from where it is deserialized in the parent process.

`bettermc4` does not involve any POSIX shared memory and hence is equivalent to `parallel`, i.e. the original `X` is serialized and transferred to the parent process using pipes.

### Character Compression
In practice, character vectors often contain a substantial amount of duplicated values.
This is exploited by `bettermc` to speed up the returning of larger character vectors from child processes:
```{r}
X <- rep(as.character(runif(1e6)), 30)
f <- function(i) X
microbenchmark::microbenchmark(
  bettermc1 = bettermc::mclapply(1:2, f),
  parallel =  parallel::mclapply(1:2, f),
  times = 1, setup = gc()
)
```

By default, `bettermc` replaces character vectors with objects of type `char_map` before returning them to the parent process:
```{r}
X_comp <- bettermc:::compress_chars(X)
str(X_comp)
```

The important detail here is the length of the `chars` vector, which just contains the unique elements of `X` and hence is significantly faster to (de)serialize than the original vector. The parent process can recover the original character vectors:

```{r}
Y <- bettermc:::uncompress_chars(X_comp)
stopifnot(identical(X, Y))
```

The internal functions `compress_chars()` and `uncompress_chars()` recursively walk the return value and apply the exported functions `char_map()` and `map2char()`, respectively.

`char_map()` is implemented using a radix sort, which makes it very efficient:
```{r}
microbenchmark::microbenchmark(
  char_map = bettermc::char_map(X),
  match = {chars <- unique(X); idx <- match(X, chars)},
  times = 3, setup = gc()
)
```
